Reasoning Graphically about Quantum Computation

### 2008-02-27 papers/reasoning-graphically/paper.tex
fixed latex bug
git-svn-id: https://isaplanner.svn.sourceforge.net/svnroot/isaplanner/trunk/quantomatic@1379 f8f375fb-4f10-0410-8d47-b6fbdc7ca053
more changes
git-svn-id: https://isaplanner.svn.sourceforge.net/svnroot/isaplanner/trunk/quantomatic@1380 f8f375fb-4f10-0410-8d47-b6fbdc7ca053

Variable nodes generalise the concept of a boundary nodes. Unlike
boundary-nodes in the concrete graph, they are not classified into
domain and co-domain. The intuition of a variable node is that it can
be replaced by any subgraph of the right type. Thus, unlike nodes in
conrete a graph, variable nodes do not define an ordering on the edges
that are incident. Instead, the edges that touch a variable node are
simply a set. The type of a variable nodes is thus defined as a
multiset rather than a tensor. Any tensor type can be lifted to a
multiset type in the obvious manner, which we denote by the operation
{\tt mset\_of\_tensor}. For example, {\tt mset\_of\_tensor($a \tensor
  b \tensor a$) = $\{(a,2), (b,1)\}$}. We say that a tensor-type, $T$,
is an instance of a multiset type, $M$, when it's lifting into a
multiset is a super-set of $M$: $ M \subseteq
\mathtt{mset\_of\_tensor}(T) \Longrightarrow T \in M$. The intuition
here is that an instance of a variable node is allowed to have as many
extra edges as it likes as long as it at least has the ones mentioned
in the original graph.

A variable node, $a$ is be said to an instance of another variable
node $b$ when the multiset representing the type of $a$ is a superset
of the multiset for the type of $b$. Thus, with regard to the
underlying semantics, $a$ being a subtype of $b$ means that the set of
instances of $a$ are a subset of the instances of $b$.

@@

@@

\begin{definition}
  \emph{merge} combines two !-boxes, $B_1$ and $B_2$ into a single
  larger !-box $B_1 \union B_2$.
\end{definition}

@@

The representation of Compact Closed Categories as Graphs, discuseed
in the previous section~\S\ref{graph-repr-comp}, is too restrictive
for reasoning about quantum computation. In particular, pictures such
the Spider Law, in Figure~\ref{spider-law-fig} are frequently
needed. Such pictures apply to an infinite family of concrete
graphs. In this section we combine {\em !-boxes} with {\em variable
  nodes} to form a representation of graphs that allows us to finitely
express certain infinite families of concrete graphs. We call this
extension \ref{graph patterns}. Using this representation Spider Law
can now be represented by the equation on graph paterns shown in
Figure~\ref{graph-pattern-spider-law-fig}. Graph matching for pattens
graphs is defined as is a suitable extension to plugging. This
provides the foundational machinery for a rewriting system, based on
pattern graphs, for reasoning about quantum computation.

>>>>

Variable nodes generalise the concept of a boundary nodes. Unlike
boundary-nodes in the concrete graph, they are not classified into
domain and co-domain. The intuition of a variable node is that it can
be replaced by any subgraph of the right type. Thus, unlike nodes in
conrete a graph, variable nodes do not define an ordering on the edges
that are incident. Instead, the edges that touch a variable node are
simply a set. The type of a variable nodes is thus defined as a
multiset rather than a tensor. Any tensor type can be lifted to a
multiset type in the obvious manner, which we denote by the operation
{\tt mset\_of\_tensor}. For example, {\tt mset\_of\_tensor($a \tensor
  b \tensor a$) = $\{(a,2), (b,1)\}$}. We say that a tensor-type $T$
is an instance of a multiset type $M$, written $T \in M$, when $M
\subseteq \mathtt{mset\_of\_tensor}(T)$. The intuition here is that
when a variable node is instantiated to some other node, it's type
will have at least the currently attached edges.

@@

%% not needed for this story...
%A variable node, $a$ is be said to an instance of another variable
%node $b$ when the multiset representing the type of $a$ is a superset
%of the multiset for the type of $b$. Thus, with regard to the
%underlying semantics, $a$ being a subtype of $b$ means that the set of
%instances of $a$ are a subset of the instances of $b$.

@@

\begin{definition}
  \emph{merge} combines two !-boxes, $B_1$ and $B_2$ into a single
  larger !-box $B_1 \cup B_2$.
\end{definition}

@@

The representation of Compact Closed Categories as Graphs, discussed
in~\S\ref{graph-repr-comp}, is too restrictive for reasoning about
quantum computation. In particular, pictures such the Spider Law, in
Figure~\ref{spider-law-fig} are frequently needed, but not
expressible. Such pictures are used to denote an infinite family of
concrete graphs that can be rewritten. In this section we combine the
{\em !-boxes} with {\em variable nodes} graph extensions to form a
representation that allows us to finitely express these infinite
families of concrete graphs. We call this extension \ref{graph
  patterns}. The Spider Law can now be represented by the equation in
Figure~\ref{graph-pattern-spider-law-fig}. We also extend the notion
of matching for graph pattens as well as present a corresponding
notion of plugging. This provides the foundations for the rewriting
machinery in~\S\ref{rewriting} which can then be used to reasoning
about quantum computation.
